---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# readxl

[![Travis-CI Build Status](https://travis-ci.org/tidyverse/readxl.svg?branch=master)](https://travis-ci.org/tidyverse/readxl) [![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/tidyverse/readxl?branch=master&svg=true)](https://ci.appveyor.com/project/tidyverse/readxl) [![Coverage Status](https://img.shields.io/codecov/c/github/tidyverse/readxl/master.svg)](https://codecov.io/github/tidyverse/readxl?branch=master) [![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/readxl)](https://cran.r-project.org/package=readxl)

The readxl package makes it easy to get data out of Excel and into R. Compared to many of the existing packages (e.g. gdata, xlsx, xlsReadWrite) readxl has no external dependencies, so it's easy to install and use on all operating systems.  It is designed to work with _tabular_ data stored in a single sheet.

readxl supports both the legacy `.xls` format and the modern xml-based `.xlsx` format. The `.xls` [libxls](http://sourceforge.net/projects/libxls/) C library is used to support `.xls`, which abstracts away many of the complexities of the underlying binary format. To parse `.xlsx`, we use the [RapidXML](http://rapidxml.sourceforge.net) C++ library.

## Installation

The easiest way to install the latest released version from CRAN is to install the whole tidyverse (NOTE: you will still need to load readxl explicitly, i.e. it is not a core tidyverse package loaded via `library(tidyverse)`.)

```{r eval = FALSE}
install.packages("tidyverse")
```

Alternatively, install just readxl from CRAN:

```{r eval = FALSE}
install.packages("readxl")
```

Or install the development version from GitHub:
 
```{r eval = FALSE}
# install.packages("devtools")
devtools::install_github("tidyverse/readxl")
```

## Usage

readxl includes example files `datasets.xlsx` and `datasets.xls`, containing datasets you already know and love. These files were created with the help of openxlsx and Excel. Paths to all example files are available via the helper `readxl_example()`.

```{r eval = FALSE}
l <- list("iris" = iris, "mtcars" = mtcars, chickwts = chickwts, quakes = quakes)
openxlsx::write.xlsx(l, file = "inst/extdata/datasets.xlsx")
```

`read_excel()` reads both xls and xlsx files and detects the format from the extension.

```{r include = FALSE}
options(tibble.print_min = 3, tibble.print_max = 3)
```

```{r}
library(readxl)

xlsx_example <- readxl_example("datasets.xlsx")
read_excel(xlsx_example)

xls_example <- readxl_example("datasets.xls")
read_excel(xls_example)
```

List the sheet names with `excel_sheets()`.

```{r}
excel_sheets(xlsx_example)
```

Specify a worksheet by name or number.

```{r}
read_excel(xlsx_example, sheet = "chickwts")
read_excel(xls_example, sheet = 4)
```

If `NA`s are represented by something other than blank cells, set the `na` argument.

```{r}
read_excel(xlsx_example, na = "setosa")
```

If you are new to the tidyverse conventions for data import, you may want to consult the [data import chapter](http://r4ds.had.co.nz/data-import.html) in R for Data Science. readxl will become increasingly consistent with other packages, such as [readr](http://readr.tidyverse.org).

## Features

* Re-encodes non-ASCII characters to UTF-8.

* Loads datetimes into POSIXct columns. Both Windows (1900) and Mac (1904) 
  date specifications are processed correctly.

* Blank columns are automatically dropped (*but this is changing!*). Blank rows that appear before the data are automatically dropped; embedded blank rows are not.

* It returns a tibble, i.e. a data frame with an additional `tbl_df` class. Among other things, this provide nicer printing.
