---
title: "Multiple Header Rows"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Multiple Header Rows}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(tibble.print_min = 4, tibble.print_max = 4)
```

```{r setup}
library(magrittr)
library(readxl)
```

A common problem when reading data into R is having multiple header rows in the source excel file. Take a look at this messy excel file from sheet 2 in the `clippy.xlsx` example file, which ships with readxl:

```{r}
readxl_example("clippy.xlsx") %>% 
    read_excel(sheet = 2)
```

**Problem:** The column names are right, but the first row of data is actually not data- it is additional metadata. This row of metadata is also causing all the columns to import as character. In reality, `death` is a datetime and `weight` is numeric.

**Solution:** We can use the `read_excel()` function to read in the same file twice. In Step 1, we’ll create a character vector of the column names only. In Step 2, we’ll read in the actual data and skip the multiple header rows at the top. When we do this, we lose the column names, so we use the character vector of column names we created in Step 1 instead.

## Step 1

In this step, we read in the first row only (by setting `n_max = 0`), extract the names from that row (using the `names()` function), and assign those to a character vector called `cnames`. This object now contains the correct column names that we’ll need in Step 2.

```{r}
(cnames <- readxl_example("clippy.xlsx") %>% 
    read_excel(sheet = 2, n_max = 0) %>% 
    names())
```

## Step 2

Now we'll read in all the rows except for the first two rows (using `skip = 2`), which contained the variable names and variable descriptions, and set the column names to `cnames`, which we created in Step 1. Now our column types are guessed correctly.

```{r}
(clippy2 <- readxl_example("clippy.xlsx") %>% 
  read_excel(sheet = 2, skip = 2, col_names = cnames))
```


If you want a way to save that metadata without polluting your actual data, you can do a third read using `n_max = 1`:

```{r}
(clippy_meta <- readxl_example("clippy.xlsx") %>% 
  read_excel(sheet = 2, n_max = 1))
```

